---
title: "Bash example"
parent: Getting started
output:
  md_document:
    variant: markdown_github
    preserve_yaml : true
---

```{r init, include=FALSE}
library(tidyverse)

source("../../R/md_helpers.R")

# change wd to where the files are located
knitr::opts_knit$set(root.dir = "examples/toys/hello_world/")
```

This vignette demonstrates how to wrap a simple 'hello world' bash script with viash.

## Demonstration
Given the files and meta files in the [examples/hello_world](examples/hello_world) directory, we demonstrate the functionality of viash in running the component 
with multiple backends.

#### Run the component
By running the component, it will output "Hello world!", followed by any other inputs provided to it.

```{bash, eval=FALSE}
cd examples/hello_world
viash run -f functionality.yaml -- I am viash!
```
```{bash, echo=FALSE}
viash run -f functionality.yaml -- I am viash!
```

#### Run the component with a Docker backend
It can also be run with a Docker backend, by providing viash with a platform yaml.
```{bash}
viash run -f functionality.yaml -p platform_docker.yaml -- General Kenobi. --greeter="Hello there."
```

#### Export as an executable
Now that we know what the component does, we can export the functionality as an executable.
```{bash}
viash export -f functionality.yaml -p platform_docker.yaml -o output
output/hello_world And now, as an executable.
```

#### viash automatically generates a CLI
By running the command with a `--help` flag, more information about the component is provided.
```{bash}
output/hello_world --help
```

#### viash allows testing the component
To verify that the component works, use `viash test`. This can be run both with or without the Docker backend.
```{bash}
viash test -f functionality.yaml -p platform_docker.yaml
```

## Developing a new component

The first step of developing this component, is writing the core functionality of the component, in this case a bash script. 

#### Write a script in Bash
This is a simple script which prints a simple message, along with any input provided to it through the `par_input` parameter.
Optionally, you can override the greeter with `par_greeter`.

```{r, echo = FALSE}
print_file("hello_world.sh")
```

Anything between the `## VIASH START` and `## VIASH END` lines will automatically be replaced 
at runtime with parameter values from the CLI. Anything between these two lines can be used to 
test the script without viash:

```{bash}
./hello_world.sh
```

Next, we write a meta-file describing the functionality of this component in YAML format.

#### Describe the functionality with a functionality.yaml

This file describes the general functionality of the component, its inputs, outputs and arguments, which 
resources are required to run it, as well as any test scripts to verify whether the component
works correctly. For each argument, you must specify the name and type, and optionally provide a description,
a default value, and a few more settings.

```{r, echo = FALSE}
print_file("functionality.yaml", "yaml")
```

For more information regarding the functionality YAML, see [functionality.md](functionality.md).

#### Describe the platform it will be run on with a platform_docker.yaml

If no platform yaml is provided, viash will execute the component natively on the host environment. 
In order to run the component with a Docker backend, the platform yaml needs to be provided. 
This file describes the dependencies of the component, that is, the base docker image and if-need-be
extra dependencies such as R or Python packages.

This component requires very little requirements, so the file is very simple.

```{r, echo = FALSE}
printFile("platform_docker.yaml", "yaml")
```

For more information regarding the platform YAML, see [platform.md](platform.md).

### Writing a first unit test
Finally, writing a unit test for a viash component is relatively simple. 
You just need to write a Bash script (or R, or Python) which runs the executable multiple
times, and verifies the output. Take note that the test needs to produce an error code not equal to
0 when a mistake is found.


```{r, echo = FALSE}
print_file("test_hello_world.sh")
```

When running the test, viash will automatically build an executable and place it -- along with other 
resources and test resources -- in a temporary working directory.

