---
layout: default
title: Config
nav_order: 4
has_children: true
---

```{r init, include=FALSE}
library(tidyverse)

source("../../R/md_helpers.R")
```

# The viash config file

A viash config file describes the functionality of the component (e.g. arguments, extra resources),
as well as the platform which will execute the component (e.g. native, docker, nextflow).
Usually, the config file is accompanied by a script which contains the actual code for the
component.

Only a small example of a viash config file is shown below, but check out the more detailed documentation regarding 
the [functionality](functionality), [native platform](platform-native), [docker platform](platform-docker), and [nextflow platform](platform-nextflow) for the full specifications for each of these subsections.

```{r, echo = FALSE}
print_file("../../examples/standalone/addrowlines/config.vsh.yaml", "config.vsh.yaml", format = "yaml")
```

```{r, echo = FALSE}
print_file("../../examples/standalone/addrowlines/script.sh", "script.sh", format = "bash")
```

The output of simply running this component.
```{bash, eval=FALSE}
viash run config.vsh.yaml -- --help
```
```{bash, echo=FALSE}
cd ../../examples/standalone/addrowlines/
viash run config.vsh.yaml -- --help
```

```{bash, eval=FALSE}
viash run config.vsh.yaml -- --input config.vsh.yaml
```
```{bash, echo=FALSE}
cd ../../examples/standalone/addrowlines/
viash run config.vsh.yaml -- --input config.vsh.yaml
```

See the [commands documentation](../../commands) for more information on the different viash commands.
