---
layout: default
title: Config
nav_order: 4
has_children: true
---

```{r init, include=FALSE}
library(tidyverse)

source("../src/helper/md_helpers.R")
```

# The viash config file

A viash config file describes the behaviour of a script and the platform it runs on.
It consists of two main sections: `functionality` and `platforms`.

The [functionality](functionality) section describes the core functionality of the component, such as 
its inputs, outputs, arguments, and extra resources. For each of the arguments, specifying
a description and a set of argument restrictions help create a useful command-line interface.
To ensure that your component works as expected, writing one or more tests is essential.

The platforms section specifies the requirements to execute the component on zero or more platforms.
The list of currently supported platforms are [Native](platform-native), [Docker](platform-docker),
and [Nextflow](platform-nextflow). If no platforms are specified, a native platform with no 
system requirements is assumed.

Usually, the config file is accompanied by a script which contains the actual code for the
component.

Only a small example of a viash config file is shown below, but check out the more detailed 
documentation regarding the [functionality](functionality), the [Native platform](platform-native), 
the [Docker platform](platform-docker), and the [Nextflow platform](platform-nextflow) for 
the full specifications for each of these subsections.

## Example

```{r, echo = FALSE}
print_file("../examples/addrowlines/config.vsh.yaml", "config.vsh.yaml", format = "yaml")
```

```{r, echo = FALSE}
print_file("../examples/addrowlines/script.sh", "script.sh", format = "bash")
```

The component can be executed as follows.
```{bash, eval=FALSE}
viash run config.vsh.yaml -- config.vsh.yaml | head -5
```
```{bash, echo=FALSE}
cd ../examples/addrowlines/
viash run config.vsh.yaml -- config.vsh.yaml | head -5
```

Would you rather use Docker as a backend? Easy as pie!
```{bash, eval=FALSE}
viash run -P docker config.vsh.yaml -- config.vsh.yaml | head -5
```
```{bash, echo=FALSE}
cd ../examples/addrowlines/
viash run -P docker config.vsh.yaml -- config.vsh.yaml | head -5
```


```{bash, eval=FALSE}
viash run config.vsh.yaml -- --help
```
```{bash, echo=FALSE}
cd ../examples/addrowlines/
viash run config.vsh.yaml -- --help
```
